---
title: "Report - Exam Group 7"
author: "Elaheh Javadi Arjmand, Ane Helland, Sigurd Mydske, Stand Hiestand"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(GGally)
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
myData <- read_delim(here("data", "dataset-exam-group7.txt"))
myData_columns <- read_delim(here("data", "dataset-exam-group7-columns.txt"))
```


## Data Source

Amy S. Nowacki, ‘Laryngoscope Dataset’, TSHS Resources Portal (2017). Available at https://www.causeweb.org/tshs/laryngoscope/

Data and description taken from [medicaldata webpage](https://higgi13425.github.io/medicaldata/reference/laryngoscope.html).

This data set contains 99 adult patients with a body mass index between 30 and 50 kg/m2 who required orotracheal intubation for elective surgery. Patient demographics, airway assessment data, intubation success rate, time to intubation, ease of intubation, and occurrence of complications were recorded.


## Working progress

This report will provide a day-to-day report of the work progress by our group. 


#### Day 5

On this day, we created an initial script which would read the exam dataset described over, and started to explore it using the head, tail and glimpse-functions. We also removed duplicate entries. We also separated some values containing multiple points of information, and cleaned up the name of the columns.
The data at the end of the day looked like this:

```{r, echo=FALSE}
glimpse(myData)
```
It obviously needed further wrangling to enable efficient analysis. 

#### Day 6

On day 6, a second dataset was joined to the original dataset using inner_join. We also changed many of the variables to factors. 

```{r, echo=FALSE}
glimpse(myData_columns)
```
Here we have wrangled several additional variables from the original dataset. 

#### Day 7

We experimented with graphical visualization of our data using ggplot. Here is an example of a geom_point with and without a regression line. 

```{r, echo=FALSE}
  plot(myData_columns %>% 
    ggplot(aes(x = bmi, y = total_intubation_time)) +
    geom_point())

```
```{r, echo=FALSE}
plot(
  myData_columns %>% 
    ggplot(aes(x = bmi, y = total_intubation_time)) +
    geom_point() + 
    geom_smooth(method = "lm"))
```

This shows the correlation between BMI and Total intubation time. 

#### Day 8
We started to use simple statistical analysis, for example using a t-test for difference in means in gender groups
```{r}
t_test_result <-
  myData_columns %>%
    t.test(total_intubation_time~gender, data=.)
print(t_test_result)
```

p-value= 0.0063, there is a difference in mean total intubation time between gender groups. 
mean total intubation time for males= 50.22 and for females= 33.68


## Day 7 & 8 Questions
Question 1: Are there any correlated measurements?
Correlation plot indicates yes.
```{r, echo=FALSE}
  ggcorr(myData_columns)

```

Question 2: Does the age distribution depend on `gender`?
It looks like there is a difference in age distribution in females vs males.
```{r, echo=FALSE}
  ggplot(data = myData_columns) +
  aes(
    x = id,
    y = age
  ) +
  geom_point() +
  facet_wrap(facets = vars(gender)) +
  geom_smooth(method = "lm")

```

Question 3: Does BMI depend on `age`?
There is a weak negative correlation based on the linear regression model. 
```{r, echo=FALSE}
 ggplot(data = myData_columns) +
  aes(
    x = age,
    y = bmi
  ) +
  geom_point()

ggplot(myData_columns, aes(x = age, y = bmi)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +  # Adds a linear regression line
  labs(title = "BMI vs Age",
       x = "Age",
       y = "BMI") +
  theme_minimal()

```

Question 4: Do age and BMI have a linear relationship?
There appears to be a weak, negative correlation.
```{r, echo=FALSE}
  ggplot(myData_columns,
         aes(
           x=age, 
           y=bmi
         ))+
  geom_point() +
  geom_smooth(method = "lm")

```

Question 5: Does number of failures change with age?
The plot may give the impression that there are a higher number of failures in older age groups
```{r, include=FALSE}
 dataset_tidy <- read_delim(here("data", "dataset-exam-group7-columns.txt"))
  
dataset_tidy <- dataset_tidy %>%
    mutate(age_cat = as.factor(case_when(age < 30 ~ "20-30", 
                                         age > 30 & age < 40 ~ "30-40",
                                         age > 40 & age < 50 ~ "40-50",
                                         age > 50 & age < 60 ~ "50-60",
                                         age > 60 & age < 70 ~ "60-70",
                                         age > 70 & age < 80 ~ "70-80"
    )))
levels(dataset_tidy$age_cat)
  
dataset_tidy %>% count(age_cat)
  
fct_drop("NA")

```
 
    
```{r, echo=FALSE}
    ggplot(data = dataset_tidy %>% filter(!is.na(age))) +
    aes(
      x = failures
    ) +
    geom_bar(
      size = 2,
    )+
    facet_wrap(vars(age_cat)
    ) 
```
  
Question 6: Does the randomzation depend on the gender?
The p-value of 1 indicated no dependency between gender and randomization
```{r, echo=FALSE}
table_data <- table(dataset_tidy$gender, dataset_tidy$randomization)
  chi_test <- chisq.test(table_data)
  print(chi_test)
```
  
Question 7: Does the randomization depend on BMI?
The p value is 0.49, not significant. Randomization does not depend on bmi.
```{r, echo=FALSE}
myData_columns %>% 
  t.test(bmi~randomization, data = .)
``` 
 
Question 8: Is there an association between BMI and total intubation time?
From the regression model, there seems to be a correlation of decreasing intubation time with increasing bmi
```{r, echo=FALSE}
 #regression (plotting)
  
  myData_columns %>% 
    ggplot(aes(x = bmi, y = total_intubation_time)) +
    geom_point() + 
    geom_smooth(method = "lm")
  #line of best fit, there appears to be a negative association between bmi and 
  #total intubation time

  

``` 

Question 9: According to the data, was there a difference of total intubation time between gender categories?
P-value= 0.0063, there is a difference in mean total intubation time between gender groups. Mean total intubation time for males= 50.22 and for females= 33.68
```{r, echo=FALSE}
 dataset_tidy %>%
    t.test(total_intubation_time~gender, data=.)
``` 


### Source material
```{r, echo=FALSE}
sessionInfo()
```

